Manuscript results: Forecasting influenza using ARIMA and ARIMAX

### 2025-05-06 manuscript_results.Rmd
update

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales")  # For `breaks_extended`
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
```

@@

    ES27_AR = (-(1 - (data$ES27_AR/data$AUTO_AR)) * 100),
    ES64_AR = (-(1 - (data$ES64_AR/data$AUTO_AR )) * 100),

    AUTO_ADJ = (-(1 - (data$AUTO_ADJ/data$AUTO_AR)) * 100),
    ES27_ADJ = (-(1 - (data$ES27_ADJ/data$AUTO_AR)) * 100),
    ES64_ADJ = (-(1 - (data$ES64_ADJ/data$AUTO_AR)) * 100),

    AUTO_TMP = (-(1 - (data$AUTO_TMP/data$AUTO_AR)) * 100),
    ES27_TMP = (-(1 - (data$ES27_TMP/data$AUTO_AR)) * 100),
    ES64_TMP = (-(1 - (data$ES64_TMP/data$AUTO_AR)) * 100),

    AUTO_EPI = (-(1 - (data$AUTO_EPI/data$AUTO_AR)) * 100),
    ES27_EPI = (-(1 - (data$ES27_EPI/data$AUTO_AR)) * 100),
    ES64_EPI = (-(1 - (data$ES64_EPI/data$AUTO_AR)) * 100),

    AUTO_AVG = (-(1 - (data$AUTO_AVG/data$AUTO_AR)) * 100),
    ES27_AVG = (-(1 - (data$ES27_AVG/data$AUTO_AR)) * 100),
    ES64_AVG = (-(1 - (data$ES64_AVG/data$AUTO_AR)) * 100),

    AUTO_AR_LB = (-(1 - (data$AUTO_AR_LB/data$AUTO_AR)) * 100),
    ES27_AR_LB = (-(1 - (data$ES27_AR_LB/data$AUTO_AR)) * 100),
    ES64_AR_LB = (-(1 - (data$ES64_AR_LB/data$AUTO_AR)) * 100),

    AUTO_ADJ_LB = (-(1 - (data$AUTO_ADJ_LB/data$AUTO_AR)) * 100),
    ES27_ADJ_LB = (-(1 - (data$ES27_ADJ_LB/data$AUTO_AR)) * 100),
    ES64_ADJ_LB = (-(1 - (data$ES64_ADJ_LB/data$AUTO_AR)) * 100),

    AUTO_TMP_LB = (-(1 - (data$AUTO_TMP_LB/data$AUTO_AR)) * 100),
    ES27_TMP_LB = (-(1 - (data$ES27_TMP_LB/data$AUTO_AR)) * 100),
    ES64_TMP_LB = (-(1 - (data$ES64_TMP_LB/data$AUTO_AR)) * 100),

    AUTO_EPI_LB = (-(1 - (data$AUTO_EPI_LB/data$AUTO_AR)) * 100),
    ES27_EPI_LB = (-(1 - (data$ES27_EPI_LB/data$AUTO_AR)) * 100),
    ES64_EPI_LB = (-(1 - (data$ES64_EPI_LB/data$AUTO_AR)) * 100),

    AUTO_AVG_LB = (-(1 - (data$AUTO_AVG_LB/data$AUTO_AR)) * 100),
    ES27_AVG_LB = (-(1 - (data$ES27_AVG_LB/data$AUTO_AR)) * 100),
    ES64_AVG_LB = (-(1 - (data$ES64_AVG_LB/data$AUTO_AR)) * 100)    
  ))
}

@@

# Heatmap plot with models ordered by WeekAhead == 1  
plot_m_heatmap <- ggplot(p_df, aes(x = factor(WeekAhead), y = Model, fill = m)) +
  geom_tile(aes(fill = ifelse(Significance == "Significant", m, NA)), color = "black") +  # Fill only significant ones
  geom_tile(data = p_df %>% filter(Significance == "Not Significant"), 
            aes(x = factor(WeekAhead), y = Model), fill = "gray80", color = "black") +  # Gray for non-significant
  geom_text(aes(label = paste0("m:", round(m, 2),", sd:", round(sd, 2))), 
            color = "black", size = 4) +  # Add text labels
  scale_fill_gradient(low = "cyan", high = "red", na.value = "gray80", 
                      name = "Percentage (%)", limits = c(-35, 35)) +  # Fixed color scale from -30 to 30
  labs(title = "Figure X: Mean differences between models' WIS and the baseline across 48 states",
       subtitle = "Gray boxes indicate models that are not significantly different from the baseline (p>0.05)",
         caption = "m = mean difference (%), sd = standard deviation",
       x = "Week Ahead",
       y = "") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
  plot.caption = element_text(size = 12),
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),) +
  facet_wrap(~ Model_Type, scales = "free_y")  # Separate NoLg and other models

# Print plot
print(plot_m_heatmap)
ggsave("Fig3.jpg", plot_m_heatmap, width = 15, height = 7 )

@@

```{r}
# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("Mean WIS improvement in percentage (W1)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
    subtitle = "(1-(Baseline/AUTO_AVG_LB))*100"
            , 
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_1WEEK<-ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_1WEEK

@@

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("Mean WIS improvement in percentage (W2)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
    subtitle = "(1-(Baseline/AUTO_AVG_LB))*100"
            , 
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_2WEEK

@@

ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("Mean WIS improvement in percentage (W3)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
    subtitle = "(1-(Baseline/AUTO_AVG_LB))*100"
            , 
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_3WEEK

@@

ES_4WEEK <- ggplot(W4_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("Mean WIS improvement in percentage (W4)") +
  theme_light() +
  theme(legend.position = "right") +
  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 3,
               color = "black",
               check_overlap = TRUE) +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
    subtitle = "(1-(Baseline/AUTO_AVG_LB))*100"
            , 
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_4WEEK

@@

library(patchwork)

# Ensure each plot has coord_sf applied
ES_1WEEK <- ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_2WEEK <- ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_3WEEK <- ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_4WEEK <- ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

# Combine in a 2x2 grid
combined_plot <- (ES_1WEEK | ES_2WEEK) /
                 (ES_3WEEK | ES_4WEEK)

# Print
print(combined_plot)

# Save to file
ggsave("Fig5_2x2_combined.jpg", combined_plot, width = 15, height = 8)
```

>>>>

```{r}
library("tidyr")
library("MMWRweek")
library("data.table")
library("caret")
library("purrr")
library("dplyr")
library("tseries")
library("gtools")
library("forecast")
library("scoringutils")
library("covidHubUtils")
library("parallel")
library("future")#https://cran.r-project.org/web/packages/future/vignettes/future-4-issues.html
library("listenv")
library("epitools")
library("ggplot2")
library("sf")
library("forcats")
library("ggplot2")
library("sf")
library("dplyr")
library("scales")  # For `breaks_extended`
library("ggplot2")
library("broom")
library("fields")
library("ggpubr")
library("patchwork")

@@

    ES27_AR = (((data$ES27_AR/data$AUTO_AR)-1) * 100),
    ES64_AR = (((data$ES64_AR/data$AUTO_AR )-1) * 100),

    AUTO_ADJ = (((data$AUTO_ADJ/data$AUTO_AR)-1) * 100),
    ES27_ADJ = (((data$ES27_ADJ/data$AUTO_AR)-1) * 100),
    ES64_ADJ = (((data$ES64_ADJ/data$AUTO_AR)-1) * 100),

    AUTO_TMP = (((data$AUTO_TMP/data$AUTO_AR)-1) * 100),
    ES27_TMP = (((data$ES27_TMP/data$AUTO_AR)-1) * 100),
    ES64_TMP = (((data$ES64_TMP/data$AUTO_AR)-1) * 100),

    AUTO_EPI = (((data$AUTO_EPI/data$AUTO_AR)-1) * 100),
    ES27_EPI = (((data$ES27_EPI/data$AUTO_AR)-1) * 100),
    ES64_EPI = (((data$ES64_EPI/data$AUTO_AR)-1) * 100),

    AUTO_AVG = (((data$AUTO_AVG/data$AUTO_AR)-1) * 100),
    ES27_AVG = (((data$ES27_AVG/data$AUTO_AR)-1) * 100),
    ES64_AVG = (((data$ES64_AVG/data$AUTO_AR)-1) * 100),

    AUTO_AR_LB = (((data$AUTO_AR_LB/data$AUTO_AR)-1) * 100),
    ES27_AR_LB = (((data$ES27_AR_LB/data$AUTO_AR)-1) * 100),
    ES64_AR_LB = (((data$ES64_AR_LB/data$AUTO_AR)-1) * 100),

    AUTO_ADJ_LB = (((data$AUTO_ADJ_LB/data$AUTO_AR)-1) * 100),
    ES27_ADJ_LB = (((data$ES27_ADJ_LB/data$AUTO_AR)-1) * 100),
    ES64_ADJ_LB = (((data$ES64_ADJ_LB/data$AUTO_AR)-1) * 100),

    AUTO_TMP_LB = (((data$AUTO_TMP_LB/data$AUTO_AR)-1) * 100),
    ES27_TMP_LB = (((data$ES27_TMP_LB/data$AUTO_AR)-1) * 100),
    ES64_TMP_LB = (((data$ES64_TMP_LB/data$AUTO_AR)-1) * 100),

    AUTO_EPI_LB = (((data$AUTO_EPI_LB/data$AUTO_AR)-1) * 100),
    ES27_EPI_LB = (((data$ES27_EPI_LB/data$AUTO_AR)-1) * 100),
    ES64_EPI_LB = (((data$ES64_EPI_LB/data$AUTO_AR)-1) * 100),

    AUTO_AVG_LB = (((data$AUTO_AVG_LB/data$AUTO_AR)-1) * 100),
    ES27_AVG_LB = (((data$ES27_AVG_LB/data$AUTO_AR)-1) * 100),
    ES64_AVG_LB = (((data$ES64_AVG_LB/data$AUTO_AR)-1) * 100)    
  ))
}

@@

# Heatmap plot with models ordered by WeekAhead == 1  
plot_m_heatmap <- ggplot(p_df, aes(x = factor(WeekAhead), y = Model, fill = m)) +
  geom_tile(aes(fill = ifelse(Significance == "Significant", m, NA)), color = "black") +  # Fill only significant ones
  geom_tile(data = p_df %>% filter(Significance == "Not Significant"), 
            aes(x = factor(WeekAhead), y = Model), fill = "gray80", color = "black") +  # Gray for non-significant
  geom_text(aes(label = paste0("m:", round(m, 1),", sd:", round(sd, 1))), 
            color = "black", size = 4) +  # Add text labels
  scale_fill_gradient(low = "cyan", high = "red", na.value = "gray80", 
                      name = "Percentage (%)", limits = c(-35, 35)) +  # Fixed color scale from -30 to 30
  labs(title = "Figure X: Mean differences between models' WIS and the baseline across 48 states",
       subtitle = "Gray boxes indicate models that are not significantly different from the baseline (p>0.05)",
         caption = "m = mean difference (%), sd = standard deviation",
       x = "Week Ahead",
       y = "") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
  plot.caption = element_text(size = 16),
    plot.title = element_text(size = 18, face = "bold"),
    plot.subtitle = element_text(size = 14),) +
  facet_wrap(~ Model_Type, scales = "free_y")  # Separate NoLg and other models

# Print plot
print(plot_m_heatmap)
ggsave("Fig3.jpg", plot_m_heatmap, width = 14, height = 7 )

@@

```{r}
# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_1WEEK <- ggplot(W1_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("1 Week Ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W1_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 2.5,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_1WEEK<-ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

@@

# Assuming merged_data already includes the necessary columns: model and percentage_improveme
# Create the map plot for 1 Week Ahead with best models and percentage improvement
ES_2WEEK <- ggplot(W2_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("2 Weeks Ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W2_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 2.5,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

@@

ES_3WEEK <- ggplot(W3_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("3 Weeks Ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W3_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 2.5,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

@@

ES_4WEEK <- ggplot(W4_map) +
  geom_sf(aes(fill = percentage_improvement)) +  # Fill based on the best model
  scale_fill_gradient2(low ="skyblue" , mid = "lightyellow", high = "darkred", midpoint = 0, limits = c(-60, 30)) +  
  ggtitle("4 Weeks Ahead") +
  theme_light() +
  theme(legend.position = "right",
    plot.title = element_text(hjust = 0.5)) +
  geom_sf_text(data = W4_map, aes(label = round(percentage_improvement,0)),  # Round to whole numbers
               size = 2.5,
               color = "black",
               check_overlap = TRUE,fontface = "bold") +  # Display percentage improvement in each state
  labs(
    fill = "%",  # Label for the legend
       x = "",
       y = ""
  )  # Adding a subtitle

@@

ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

@@

```{r}
# Ensure each plot has coord_sf applied
ES_1WEEK <- ES_1WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_2WEEK <- ES_2WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_3WEEK <- ES_3WEEK + coord_sf(xlim = x_limits, ylim = y_limits)
ES_4WEEK <- ES_4WEEK + coord_sf(xlim = x_limits, ylim = y_limits)

# Combine in a 2x2 grid
combined_plot <- (ES_1WEEK | ES_2WEEK) /
                 (ES_3WEEK | ES_4WEEK) + 
  plot_annotation(
    title = "Figure X: Mean WIS differences (%) compared to the baseline",
    subtitle = "Results by forecast horizon",
    theme = theme(plot.title = element_text(size = 16, face = "bold"),
                  plot.subtitle = element_text(size = 14))
  )
# Print
print(combined_plot)

# Save to file
ggsave("Fig5.jpg", combined_plot, width =12, height =7)
```
